### 2018-03-09 paper/background.tex
Finish S2.5

\noindent
The above makefile lists three tasks: (i) compile a utility library comprising
files \cmd{util.h} and \cmd{util.c} into \cmd{main.o} by
executing\footnote{In this example we treat \cmd{gcc} as a pure function for the
sake of simplicity. In reality there are multiple versions of \cmd{gcc} and the
actual binary that is used to compile and link files is also listed as a task
dependency.} the command \cmd{gcc -c util.c}, (ii) compile the main source file
\cmd{main.c} into \cmd{main.o}, and (iii) link object files \cmd{util.o} and
\cmd{main.o} into the executable \cmd{main.exe}. The makefile contains the
complete information about the \emph{task dependency graph}, which is shown in
Fig.~\ref{fig-make}(a).

@@

\Excel's recalculation algorithm~\cite{excel_recalc} on the other hand is
significantly different from \Make. Since it is impossible to construct the full
dependency graph upfront in the presence of dynamic dependencies, \Excel
maintains the \emph{calculation chain}, which is an approximation to the correct
topological order. During recalculation, \Excel processes cells in this order,
but can \emph{defer recalculation of a cell} by moving it down the chain when newly
discovered dynamic dependencies dictate that. We discuss this algorithm in more
detail in~\S\ref{sec-examples}.

@@

\Shake's implementation is different from both \Make and \Excel in two aspects.
First, it uses the dependency graph produced in the previous build to detect
which files are dirty -- a powerful idea with a long history of use in
\emph{incremental}~\cite{demers1981incremental},
\emph{adaptive}~\cite{acar2002adaptive}, and
\emph{self-adjusting}~\cite{acar2007selfadjusting} computation (see a discussion
in related work~\S\ref{sec-related}). Second, instead of abandoning and deferring
the computation of tasks whose rediscovered dependencies are dirty (as \Excel
does), \Shake \emph{pauses} their computation until the dependencies are
brought up-to-date.

@@

To support cloud builds, \Bazel maintains a conventional \emph{content-addressable
cache}, which can be used to fetch a previously built file given the hash of its
content, as well as \emph{dependency graphs from all previous builds}, annotated
with observed file hashes\footnote{Here we ignore the issue of limited cloud
storage resources for the sake of simplicity; in practice, old entries are
regularly evicted from the storage, as further discussed in~\S\ref{sec-engineering}.}.
The latter allows to bypass the execution of a task, by predicting the hash of
its result from the hashes of the dependencies, and subsequently fetch the file
from the cache. A concrete implementation is provided in~\S\ref{sec-examples}.

\subsection{Overview}
\label{sec-background-overview}

\begin{table}[t]
\small
\centering
\begin{tabular}{l||l|l|l|l||l|l}
Build system & Dependencies & Minimal & Early cutoff & Cloud & Metadata & Algorithm \\
\hline
\Make      & Static  & Yes & No  & No  & Dirty bit        & Topological   \\
\Excel     & Dynamic & No  & No  & No  & Dirty bit, chain & Chain + defer \\
\Shake     & Dynamic & Yes & Yes & No  & Graph            & Recursive     \\
\Bazel     & Static  & Yes & Yes & Yes & Graphs, cache    & Topological   \\
\hline
This paper & Dynamic & Yes & Yes & Yes & Graphs, cache    & Recursive     \\
\end{tabular}
\vspace{2mm}
\caption{Overview of build systems.\label{tab-overview}}
\label{tab:use-cases}
\end{table}

\subsection{Abstract vocabulary for build systems}
\label{sec-background-vocabulary}

@@

    \item Most build systems persistently store auxiliary \emph{build
    information} for profiling and optimisation purposes: \Make stores file
    modification times, \Shake stores the discovered dynamic dependency graph,
    \Bazel and other cloud build systems store information about inputs and
    outputs of previously executed tasks, etc.
\end{itemize}

>>>>

\noindent
The above makefile lists three tasks: (i) compile a utility library comprising
files \cmd{util.h} and \cmd{util.c} into \cmd{util.o} by
executing\footnote{In this example we treat \cmd{gcc} as a pure function for the
sake of simplicity. In reality, there are multiple versions of \cmd{gcc} and the
actual binary that is used to compile and link files is often also listed as a task
dependency.} the command \cmd{gcc -c util.c}, (ii) compile the main source file
\cmd{main.c} into \cmd{main.o}, and (iii) link object files \cmd{util.o} and
\cmd{main.o} into the executable \cmd{main.exe}. The makefile contains the
complete information about the \emph{task dependency graph}, which is shown in
Fig.~\ref{fig-make}(a).

@@

\Excel's recalculation algorithm~\cite{excel_recalc} on the other hand is
significantly different from \Make. Since it is impossible to construct the full
dependency graph upfront in the presence of dynamic dependencies, \Excel
maintains the \emph{calculation chain}, which is an approximation to the correct
topological order produced by the previous build. During recalculation, \Excel
processes cells in this order, but can \emph{defer recalculation of a cell} by
moving it down the chain when one of its newly discovered dynamic dependencies
is dirty. We discuss this algorithm in more detail in~\S\ref{sec-examples}.

@@

\Shake's implementation is different from both \Make and \Excel in two aspects.
First, it uses the dependency graph produced in the previous build to detect
which files are dirty -- a powerful idea with a long history of use in
\emph{incremental}~\cite{demers1981incremental},
\emph{adaptive}~\cite{acar2002adaptive}, and
\emph{self-adjusting}~\cite{acar2007selfadjusting} \emph{computations} (see a
discussion in related work~\S\ref{sec-related}). Second, instead of abandoning
and deferring the computation of tasks whose rediscovered dependencies are dirty
(as \Excel does), \Shake \emph{pauses} their computation until the dependencies
are brought up-to-date. We refer to this build algorithm as \emph{recursive}.

@@

To support cloud builds, \Bazel maintains a conventional \emph{content-addressable
cache}, which can be used to fetch a previously built file given the hash of its
content, as well as \emph{dependency graphs from all previous builds}, annotated
with observed file hashes\footnote{Here we ignore the issue of limited cloud
storage resources for the sake of simplicity; in practice, old entries are
regularly evicted from the storage, as further discussed in~\S\ref{sec-engineering}.}.
The latter allows to bypass the execution of a task, by predicting the hash of
its result from the hashes of the dependencies, and subsequently fetch the result
from the cache. A concrete implementation is provided in~\S\ref{sec-examples}.

\subsection{Summary}
\label{sec-background-summary}

We summarise differences between four discussed build systems in
Table~\ref{tab-summary}. The `Metadata' column refers to the \emph{auxiliary
build information} persistently stored between builds by all build systems:
\begin{itemize}
    \item \Make stores file modification times, or rather, it relies on the file
    system to do that. In principle a single dirty bit is sufficient, as will be
    demonstrated in~\S\ref{sec-examples}.
    \item \Excel stores the dirty bit as well as the calculation chain.
    \item \Shake stores the dependency graph discovered in the previous build,
    annotated with file content hashes for efficient checking of file changes.
    \item \Bazel stores all dependency graphs discovered in previous builds
    annotated with file hashes, and the content-addressable cache.
\end{itemize}

The last row of the table describes the new build system proposed in this paper,
which shares the advantages of \Shake and \Bazel.

\begin{table}[h]
\smaller
\centering
\begin{tabular}{l||l|c|c|c|c||l|l}
\hline
$\!\!$Build system$\!$&$\!$Dependencies$\!$&$\!$Minimal$\!$&$\!$Self-tracking$\!$&$\!$Cutoff$\!$&$\!$Cloud$\!$&$\!$Metadata$\!$&$\!$Algorithm$\!\!$\\\hline
$\!\!$\Make       $\!$&$\!$Static  & Yes & No  & No  & No  &$\!$Dirty bit       $\!$&$\!$Topological $\!\!$\\
$\!\!$\Excel      $\!$&$\!$Dynamic & No  & Yes & No  & No  &$\!$Dirty bit, chain$\!$&$\!$Chain, defer$\!\!$\\
$\!\!$\Shake      $\!$&$\!$Dynamic & Yes & No  & Yes & No  &$\!$Graph           $\!$&$\!$Recursive   $\!\!$\\
$\!\!$\Bazel      $\!$&$\!$Static  & Yes & No  & Yes & Yes &$\!$Graphs, cache   $\!$&$\!$Topological $\!\!$\\\hline
$\!\!$This paper  $\!$&$\!$Dynamic & Yes & No? & Yes & Yes &$\!$Graphs, cache   $\!$&$\!$Recursive   $\!\!$\\
\hline
\end{tabular}
\vspace{2mm}
\caption{Summary of build system differences.\label{tab-summary}}
\label{tab:use-cases}
\end{table}

\subsection{Common vocabulary for build systems}
\label{sec-background-vocabulary}

@@

\end{itemize}
